\input{../../tex_tools/outline_preamble.tex}

\begin{document}
\section*{Abstract/Introduction}

\section*{Relevant Background}
\begin{itemize}
	\item Griffiths experiment and Avery's identification of the
		transforming molecule. 
	\item Evolutionary consequences of inter- and intra-species gene transfer. 
	\item Distribution of competence states in bacteria - what is known and
		not known?
	\item Energetics is still a complete mystery. For the competence
		mechanisms that have been characterized, it is
		still unknown what powers the transport of DNA.
	\item Role of gene transfer in spread of antibiotic resistance and
		virulence factors. Implications for industry
		and directed evolution of large microbial communities.
\end{itemize}

\section*{Aims and Approach}
\begin{itemize} 
	\item Study the uptake of DNA at single molecule resolution.  What
		physical parameters of the DNA and the cells environment govern
		uptake of DNA? Can we generate a predictive model of DNA uptake?

	\item Monitor the segregation and partitioning of plasmids over the
		course of the cell cycle. Why does segregation of incompatible
		plasmids happen at a rate much faster than expected? In absence
		of partitioning machinery, how do plasmids faithfully partition
		into daughter cells? What role does selection play on the
		population genetics of plasmids?  

	z\item Quantify the micromechanics of plasmid dna \textit{in vivo}. How
		does this relate to the cell cycle and copy number control
		mechanisms?  

	\item Quantify the extent and spatial distribution of horizontal gene
		transfer in microbial communities such as colonies and biofilms.
		Why is gene transfer upregulated in biofilms? Does gene transfer
		serve as a robust method of signaling and communication?
		Computational simulations of planktonic cell mixtures have been
		informative, but can we apply the same methods to biofilms?  

	\item Relationship between genetic transfer and and use of DNA as a
		carbon source. Which came first and why use such a dangerous
		molecule as food?

\end{itemize}

\section*{Conclusion}



\end{document}
